---
title: "Reproducible Data Analysis of Heroes of the Storm Competitive Data"
author: "Danishswag"
output: html_notebook
---

While the purpose of data analysis should be to tell a story, this R notebook
should provide a good starting point for analyzing Heroes of the Storm (HotS)
competitive data. The idea is to make a report that I (or anyone else)
can run on a regular basis to evaluate the teams and players in the competitive
scene. Note that some matches in the `player_details` table have flipped team
names which may lead to slightly inconsistent results compared to the
Master League data. Send me a message or make a pull request if you notice
anything wrong :)

As usual, everything is under the MIT license like the rest of the repository,
so make sure to give me attribution if you use anything here.

# Setup
```{r}
require("sqldf")
require("ggplot2")
require("PlayerRatings")
require("htmlTable")
```

Setting up the connection to the SQLite database.

```{r}
db = dbConnect(SQLite(), dbname = "../hots_scrapy/HotSCompetitive.sqlite")
```


# Blizzard Statistics

On the [HGC Website](http://us.heroesofthestorm.com/esports/en/), Blizzard
often puts up [statistics from the previous weekend of games](http://us.heroesofthestorm.com/esports/en/news/20529704/stats-snapshot:-week-3).
Running the code below should give us the same information. The replay files
from Korea tend to go up on the server late, so Korean values that depend on
detailed information like hero damage are less reliable.

## Highest Average Number of Kills by Region

You can check yourself the lookup tables, but Korea's HGC ID is 37.

```{sql, connection=db}
SELECT * FROM (SELECT deets.player_name, AVG(deets.kills) AS average_kills, 
r.name AS region FROM player_details deets
JOIN match_basic basic ON deets.match_id=basic.match_id
JOIN tournaments t ON basic.tournament=t.id
JOIN region r ON t.region=r.id
WHERE basic.tournament = 37
GROUP BY deets.player_name, r.name ORDER BY AVG(deets.kills) DESC
LIMIT 1 )

UNION

SELECT * FROM (SELECT deets.player_name, AVG(deets.kills) AS average_kills,
r.name AS region FROM player_details deets
JOIN match_basic basic ON deets.match_id=basic.match_id
JOIN tournaments t ON basic.tournament=t.id
JOIN region r ON t.region=r.id
WHERE basic.tournament = 33
GROUP BY deets.player_name, r.name ORDER BY AVG(deets.kills) DESC
LIMIT 1)

UNION

SELECT * FROM (SELECT deets.player_name, AVG(deets.kills) AS average_kills,
r.name AS region FROM player_details deets
JOIN match_basic basic ON deets.match_id=basic.match_id
JOIN tournaments t ON basic.tournament=t.id
JOIN region r ON t.region=r.id
WHERE basic.tournament = 32
GROUP BY deets.player_name, r.name ORDER BY AVG(deets.kills) DESC
LIMIT 1)
```

# Team Statistics
## Glicko Ratings

We need to pull out the data for the glicko code.
```{r}
glicko.df <- sqldf("select week,
                      home_team,
                      away_team,
                      score 
                    from match_basic", conn=db)
```

In this section we are going to generate the Glicko ratings for all of the
teams in the league. This section will only output the top ten teams. In
order to actually make predictions, these ratings will have to be passed into
another script.

```{r}
glicko.ratings <- glicko(glicko.df)

# Exporting ratings and printing the top 20 teams
htmlTable(txtRound(glicko.ratings$ratings[1:20, ], 0))
glicko.ratings$ratings[1:20, ]
```

## Shortest and Longest Average Match Duration
There is only duration data for some matches, so this is certainly not a
perfect calculation. It includes data since Sep. 1st, 2016. If you want
to see the shortest durations, switch `DESC` to `ASC`.

```{sql, connection=db}
SELECT team_name, AVG(duration) AS avg_duration, COUNT(team_name) AS matches
FROM
(
SELECT home_team AS team_name, duration
  FROM match_basic
  WHERE duration IS NOT NULL
  AND DATE > '2016-09-01'
UNION ALL

SELECT away_team AS team_name, duration
  FROM match_basic
  WHERE duration IS NOT NULL
    AND date > date('now', 'start of month', '-6 months')
)
GROUP BY team_name
HAVING count(team_name) > 10
ORDER BY AVG(duration) DESC
LIMIT 10
```

## Most Banned Against
The idea here is to calculate which characters are banned the most
against a given team in the last 6 months.

```{sql, connection=db}
SELECT ban AS most_banned, COUNT(ban) AS frequency FROM (
    SELECT CASE WHEN home_team = 'Fnatic' THEN home_ban1 ELSE (
        CASE WHEN away_team = 'Fnatic' THEN away_ban1 END) END AS ban
    FROM match_basic
    WHERE home_team = 'Fnatic' OR away_team = 'Fnatic'
    
    UNION ALL
    
    SELECT CASE WHEN home_team = 'Fnatic' THEN home_ban2 ELSE (
        CASE WHEN away_team = 'Fnatic' THEN away_ban2 END) END AS ban
    FROM match_basic
    WHERE (home_team = 'Fnatic' OR away_team = 'Fnatic')
      AND date > date('now', 'start of month', '-6 months')
    )
GROUP BY ban
ORDER BY COUNT(ban) DESC LIMIT 10
```

# Player Level Statistics
## Finding the 'Hero Pool'
I define the hero pool as all characters with which someone has great than a
50% win rate with, although this is dependent on the quality of the team. To
account for this, I also allow a hero to count in a player's hero pool if
they have played them more than 10 times in competitive play.
```{sql, connection=db}
SELECT hero_name,
  COUNT(hero_name) as num_matches,
  SUM(is_win)*1.0 / COUNT(is_win) * 100 AS win_percent
  FROM
    (
    SELECT d.hero_name,
        CASE WHEN ((m.score = 0 AND m.away_team = d.team_name)
                    OR (m.score = 1 AND m.home_team = d.team_name))
            THEN 1
            ELSE 0 END AS is_win
    FROM player_details d
    JOIN match_basic m ON d.match_id = m.match_id
	  WHERE d.player_name = 'Swoy'
    )
  GROUP BY hero_name
  HAVING SUM(is_win)*1.0 / COUNT(is_win) >= 0.5
      OR COUNT(hero_name) >= 10
  ORDER BY SUM(is_win)*1.0 / COUNT(is_win) DESC
  LIMIT 10
```

## Finding the Players with the Highest Win Rates
These will mostly be based on what teams players are a part of, especially
in the early days of this competitive scene. Later on though, it may be
make it possible to discern which player is actually helping a team win.
```{sql, connection=db}
SELECT player_name,
  SUM(is_win)*1.0 / COUNT(*) AS win_rate,
  count(*) AS matches_played
  FROM
  (
    SELECT m.match_id,
        m.date,
        d.team_name,
        d.player_name,
        d.hero_name,
        CASE WHEN ((m.score = 0 AND m.away_team = d.team_name)
                    OR (m.score = 1 AND m.home_team = d.team_name))
            THEN 1
            ELSE 0 END AS is_win
    FROM player_details d
    JOIN match_basic m on d.match_id = m.match_id
  )
  GROUP BY player_name
  HAVING COUNT(*) > 30
  ORDER BY SUM(is_win)*1.0 / COUNT(*) DESC
  LIMIT 10
```

# Final Steps
Make sure to always close your connection to the database to be polite.
```{r}
dbDisconnect(db)
```
